---
title: "WHAM_Run43_diagnostic_testing"
output: html_document
date: '2022-08-01'
---
  
## Test AMO 
This run is identical to WG run 43 (fit to AMO without link to R specified) but has artificially reduced the AMO standard error to test if diagnostic issues for the environmental covariate in this run could be attributed to large standard errors. 
  
## Model description
Model specified as in run 29B with 2 major differences, 1) both age 4 and 5 were fixed at full selectivity for all indices (impacts Bigelow spring and Albatross fall) and 2) this run was fit with an environmental covariate that has no effect so it can be directly compared to run 38A that incorporates this environmental effect on recruitment. The Atlantic Multidecadal Oscillation (AMO) was the environmental covariate considered here.

### Load R packages
```{r, echo=FALSE}
library(tidyverse)
library(wham)
library(readxl)
library(DataExplorer)
```

### Load data
```{r}
# ASAP data input
asap3 <- read_asap3_dat(paste(here::here(), "data", "PlaiceWHAM-2019_revised_NEFSC-LW-WAA_splitNEFSC-BigUnit.DAT", sep="/"))

# Load and filter environmental data to match year range of other model data
AMO <- read.csv(paste(here::here(), "data", "AMO_se.csv", sep="/"))
# Add placeholder for missing years of data
AMO <- rbind(c(1981, -999, -999, -999, -999), AMO) # 1981 years in the wrong order will cause R to bomb
AMO <- rbind(c(1980, -999, -999, -999, -999), AMO) # 1980
AMO <- rbind(c(1979, -999, -999, -999, -999), AMO) # 1979
```

### Artificially decrease the scale of the standard error
This section was added to check model diagnostics for the fit to the environmental covariate.
```{r}
# Reduce scale of standard error
AMO$se_scaled <- AMO$se/2

```
Check with plot
```{r}
AMO %>%
  filter(Year > 1981) %>%
  ggplot() +
  geom_ribbon(aes(ymin = mean-se, ymax = mean+se, x=Year, alpha = 0.05)) + # Original SE
  geom_ribbon(aes(ymin = mean-se_scaled, ymax = mean+se_scaled, x=Year, alpha = 0.05)) + # Scaled SE
  geom_line(aes(x=Year, y=mean)) 
  
```

### Prepare model input for full run
```{r}
NAA_re = list(sigma = "rec+1") # Full state-space model
NAA_re$cor = "iid" # iid random effects
NAA_re$recruit_model = 2 # recruitment is random about the mean
NAA_re$recruit_pars = exp(10) #initial guess for mean recruitment

# Setup environmental covariate and options
ecov <- list(
  label = "AMO",
  mean = as.matrix(AMO$mean),
  # logsigma = as.matrix(AMO$se_scaled), # For artificially scaled down standard errors
  logsigma = as.matrix(AMO$se), # For SEs used in WG run 43
  year = as.numeric(AMO$Year),
  use_obs = matrix(c(FALSE, FALSE, FALSE, rep(TRUE, nrow(AMO)-3)), ncol = 1, nrow=nrow(AMO)), # use all obs except the first 3 (no data 1979, 1980, and 1981 not full year of data in 1981) 
  lag = 1, # AMO anomaly in year t impact R in year t+1
  #process_model = "ar1",
  process_model = "rw",
  where = "none", 
  how = 0 
)

# Setup initial selectivity model and parameters
use_n_indices = 4
modelsetup <- c(rep("logistic", asap3$dat$n_fleet_sel_blocks), rep("age-specific", use_n_indices))

init_fleet_sel <- list(c(2,0.4)) # logistic parameters, based on model type
init_index_sel <- lapply(1:use_n_indices, function(x) c(0.5, 0.5, 0.5, 1, 1, 0.5, 0.5, 0.5, 0.5, 0.5, 0.5))

# Setup fixed parameters
fix_fleet_sel <- lapply(1:asap3$dat$n_fleets, function(x) NA) 
fix_index_sel <- lapply(1:use_n_indices, function(x) NA) # Set up index object

fix_index_sel[[1]] <- c(4,5) # Fix age 4 & 5 for for index 1 (NEFSC spring Albatross)
fix_index_sel[[2]] <- c(4,5) # Fix age  5 for for index 2 (NEFSC spring Bigelow)
fix_index_sel[[3]] <- c(4,5) # Fix age 4 for for index 3 (NEFSC fall Albatross)
fix_index_sel[[4]] <- c(4,5) # Fix age 4 for for index 4 (NEFSC fall Bigelow)

# Setup random effect by selectivity block (here: fleet, index1, index2)
randeffect <- c(rep("iid", asap3$dat$n_fleet_sel_blocks), rep("iid", use_n_indices))

# Setup random effect by selectivity block (here: fleet, index1, index2, index3, index4)
randeffect <- c(rep("iid", asap3$dat$n_fleet_sel_blocks), rep("none", 4)) # Don't include selectivity random effects for any surveys 

# Setup selectivity list
sel_list <- list(model = modelsetup, # list selectivity model for each fleet and index
                 re = randeffect,
                 initial_pars = c(init_fleet_sel, init_index_sel),
                 fix_pars = c(fix_fleet_sel, fix_index_sel))

input <- prepare_wham_input(asap3, NAA_re = NAA_re, selectivity = sel_list, ecov=ecov, model_name = "Debugging") 
```

### Save input
```{r}
# Save model data input
saveRDS(input, file=here::here("PlaiceMSE", "Diagnostic_development", "Debug_R-AMO", "input_debugRAMO.rds"))
```

### Fit model, check convergence, and run diagnostics
```{r}
WHAM_debugRAMO <- fit_wham(input, do.osa = TRUE, do.retro = TRUE) 
check_convergence(WHAM_debugRAMO)
print(paste("Number of parameters", length(WHAM_debugRAMO$par), sep=" "))

plot_wham_output(mod=WHAM_debugRAMO, out.type='html', dir.main = here::here("PlaiceMSE", "Diagnostic_development", "Debug_R-AMO"))
```

### Save output
```{r}
# Save fitted model
saveRDS(WHAM_debuRAMOg, file=here::here("PlaiceMSE", "Diagnostic_development", "Debug_R-AMO", "WHAM_debugAMO_model.rds"))
```

